The Minimal Retroreflective Microfacet Model

### 2025-12-19 article.tex
Add updated figures, and improvements to text.

\setciteauthor{Matthias Raab and Laurent Belcour and Frankie Liu and Jamie Portsmouth}
\setcitetitle{The Minimal Retroreflective Microfacet Model (MRMM)}
%\setheadtitle{Abbreviated title, only if full title won't fit in page headers}

@@

@@

\teaser{
  \vspace*{-0.5cm}
  \centering
    \includegraphics[width=0.32\linewidth]{figures/sphere_side_r0.3.png     }
    \includegraphics[width=0.32\linewidth]{figures/sphere_quarter_r0.3.png  }
    \includegraphics[width=0.32\linewidth]{figures/sphere_front_r0.3.png    }
  \caption{Comparison of our retroreflector model (top) obtained by an almost trivial modification of the equivalent roughness GGX microfacet model (bottom). The lighting varies (left to right) from side-on, to quarter-on, and straight-on, exhibiting the strong retro-reflective peak.
   \vspace{-0.5cm}
  \label{fig:teaser}}
}

@@

\begin{abstract}
  \small
  We introduce the ``Minimal Retroreflective Microfacet'' (MRM) model, which provides a simple scheme for modifying a pre-existing microfacet BSDF implementation to produce a visually plausible, energy-conserving retroreflective result.
  \end{abstract}

@@

Rendering of retro-reflective materials is useful in a variety of contexts. For security garments and signs, retroreflective appearances are heavily used. A number of CG models have been proposed. For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least, b) provably energy-conserving, c) efficient and easy to implement, ideally a small modification to well-understood microfacet models. We present here a model based on the previously published ``back-vector'' formulation~\cite{BelcourBackvector} which meets these requirements. It is so simple a modification to a regular microfacet model, that we term it the ``Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible retro-reflective result.

For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least and follows measured observations, b) provably energy-conserving, c) efficient and easy to implement, ideally a small modification to well-understood surface models.
\paragraph{Previous works.}
Modeling of retro-reflectivity in the surface scattering function has received little attention so far. For example, \citet{Guo2017Retro} proposed a model to incorporate the retro-reflectivity of prismatic sheets. However, their model only provide pure retro-reflection which would lack artistic control. They later extented this model account for imperfect retro-reflection with the addition of a microfacet surface (following Beckmann's normal distribution) above the prismatic sheet~\citet{Guo2018Retro}. However, due to the coupling of retro-reflection and rough refraction, they were unable to derive a close form solution for it and relied on fitting of the ground truth model.

@@

We detail the correct usage and derivation of this retro-reflective BSDF model within the microfacet framework~\cite{WalterMicrofacet,HeitzMicrofacet}.
We first recall some basic principles from microfacet theory (Section~\ref{sec:microfacets}), then show how this model can be steered from forward- to retro-reflection (Section~\ref{sec:backvector}), and further provide proof that the resulting BSDF is fully plausible (Section~\ref{sec:plausibility}).
Finally, we sketch how the model can be implemented in practice (Section~\ref{sec:implementation}).

@@

\begin{figure}
  \begin{center}
    \begin{tabular} {|p{0.2\textwidth}|p{0.7\textwidth}|}
%    \begin{tabular} {|l|l|}
      \hline
      $\vecv$ & normalized vector to viewer \\
      \hline
      $\vecl$ & normalized vector to light \\
      \hline
      $\vecn$ & surface normal, with the convention that it points to
      the medium with lower index of refraction in cases dealing with $\vecv$ and $\vecl$ pointing to different sides\\
      \hline
      $\mathbf{t_1}, \mathbf{t_2}$ & tangent and bitangent (forming an orthonormal coordinate frame with the normal) \\
      \hline
      $\vechr$ & half vector of reflection, $\vechr = \vechr(\vecv, \vecl) = \frac{\vecv + \vecl}{\lVert \vecv + \vecl \rVert}$ \\
      \hline
      $\vecht$ & half vector of refraction, $\vecht = \vecht(\vecv, \vecl) = -\frac{\left(\etav\vecv + \etal\vecl\right)}{\lVert\etav\vecv + \etal\vecl\rVert}$, where $\etav$ and $\etal$ are the refractive indices of the media $\vecv$ and $\vecl$ point to (always points to medium with lower index of refraction)\\
      \hline
      $\vdot{\vecx}{\vecy}$ & dot product of $\vecx$ and $\vecy$ \\
      \hline
      $\text{reflect}(\mathbf{x}, \mathbf{m})$ & reflection of vector $\mathbf{x}$ on surface with normal $\mathbf{m}$, i.e. $\text{reflect}(\mathbf{x}, \mathbf{m}) = -\mathbf{x} + 2 \vdot{\mathbf{x}}{\mathbf{m}} \mathbf{m}$ \\
%      \hline
%      $\Omega$ & hemisphere of directions around the normal\\
      \hline
      $\Omega, \Omega_{\vecx}$ & hemisphere of directions around the normal, hemisphere of directions around the normal pointing to the side $\vecx$ points to\\
      \hline
    \end{tabular}
  \end{center}
  \caption{
    Notations used.
  }
  \label{fig:notation}
\end{figure}

Using the notations from Figure \ref{fig:notation}, a microfacet BRDF \cite{WalterMicrofacet, HeitzMicrofacet} takes the form (\citet{WalterMicrofacet} Equation (20))
\begin{equation}
  f_r(\vecv, \vecl) = \frac{D(\vechr) \, G_{2}(\vecv,\vecl) \, F(\vechr, \vecv)}{4 \vdot{\vecv}{\vecn} \vdot{\vecl}{\vecn}}
  \label{eq:microfacet_brdf}
\end{equation}
where $D$ is the distribution of microfacets, normalized to fulfill $\int_\Omega D(\vech) \vdot{\vecn}{\vech} d\vech = 1$, $G_{2}$ is the shadowing-masking function, and $F$ is the Fresnel term. The half-vector $\vechr$ is used as the \textit{selector} for microfacets reflecting light from incident direction $\vecl$ to outgoing direction $\vecv$ (\textit{i.e.} only microfacets with normals aligned with the half-vector will reflect light).
Similarly, using the half-vector of refraction as selector, yields the corresponding BTDF
\begin{equation}
  f_t(\vecv, \vecl) =
  \frac{\avdot{\vecv}{\vecht}\avdot{\vecl}{\vecht}\etav^2 D(\vecht) \, G_{2}(\vecv,\vecl) \, \left(1 - F(\vecht, \vecv) \right)}
       {\avdot{\vecv}{\vecn} \avdot{\vecl}{\vecn} \left(\etav \vdot{\vecv}{\vecht} + \etal \vdot{\vecl}{\vecht} \right)^2}
  \label{eq:microfacet_btdf}.
\end{equation}
In Section~4.1 and Section~4.2 of their work, Walter et al.~\cite{WalterMicrofacet} detail the effect of using the half vector as the microfacet selector and obtain Equations~\eqref{eq:microfacet_brdf} and \eqref{eq:microfacet_btdf}.
In this work, we show that we can use the back vector~\cite{BelcourBackvector} used to model retro-reflection as a microfacet selector. Before detailing the impact of such a change, we review some properties of microfacet models.

\paragraph*{Shadowing-Masking.}
The shadowing-masking function combines the visibility masking function $G_1(\vecv, \vecl)$ for view direction and
$G_1(\vecl, \vecv)$ for light direction:
\begin{equation}
G_{2}(\vecv, \vecl) = G_{2}\left(G_1(\vecv, \vecl), G_1(\vecl, \vecv) \right)
\end{equation}
Mathematically, the BSDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
$G_{2}(\vecv, \vecl) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible micro-surfaces \cite{HeitzMicrofacet}
\begin{equation}\label{eq:visible}
  \avdot{\vecv}{\vecn} = \int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) G_1(\vecv, \text{reflect}(\vecv, \vech)) \vdot{\vecv}{\vech} d\vech.
\end{equation}
Choices for $G_1$ with this property are the Smith masking function
\begin{equation}
  G_1(\vecv, \vecl) = G_{1,\text{Smith}}(\vecv) = \frac{\avdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech},
\end{equation}
where analytic expressions are available for Beckmann and GGX distributions, and the v-cavity masking function
\begin{equation}
  G_1(\vecv, \vecl) = G_{1,\text{vc}}(\vdot{\vecv}{\vecn},\vdot{\vecn}{\vech},\vdot{\vecv}{\vech}) =
  \min\left\lbrace\frac{2\avdot{\vecv}{\vecn}\avdot{\vecn}{\vech}}{\avdot{\vecv}{\vech}}, 1\right\rbrace,
\end{equation}
which is generally applicable for distributions with symmetry around the normal.

@@

As illustrated in Figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as an reflection of the view direction on a plane aligned with $\vecv$ and perpendicular to the surface.
Intuitively, this is equivalent of observing the microfacet reflection through a virtual mirror, thus turning glossy forward-reflection into glossy retro-reflection.
While there is no immediate physical foundation to this construction, this interpretation can serve as motivation why components from the regular microfacet model still work when changing the domain to the back-vector.

\paragraph*{Jacobian of Back-Vector.}
First we observe that Jacobian of the change of variable from the back-vector to the light vector $\vecl$ for specular microfacets~(Equation~(11) and~(14) of \citet{WalterMicrofacet}) has the same form as the half-vector one:
\begin{equation}
\left\lVert \dfrac{\partial \omega_{\vecb_{r}}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech_{r}(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{1}{4 \left| \vdot{\vecl}{\vech(\vecvv, \vecl)} \right|} = \dfrac{1}{4 \left| \vdot{\vecl}{\vecb} \right|}
\end{equation}
We also observe that this is true for refractive microfacets~(Equation (17) of \citet{WalterMicrofacet}):
\begin{equation}
\left\lVert \dfrac{\partial \omega_{\vecb_{t}}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech_{t}(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{  \eta_o^2 \left| \vecl, \vech_{t}(\vecvv, \vecl)  \right|}{ \left\lVert \vech_{t}(\vecvv, \vecl) \right\rVert^2}
\end{equation}
This is due to the fact that we have to differentiate with respect to the light direction (which is not altered by the back vector).

\paragraph*{Microfacet reflection term.}
There is experimental evidence~\cite{BelcourBackvector} that using a Fresnel term based on the back-vector can provide a good match to actual retro-reflective material measurements for a reflection back-vector based BRDF.
As such, we update the formulation of $\rho(\vecv, \vecm)$~(Equation (11) of \citet{WalterMicrofacet}) to use the reflected view vector instead. It follows that the reflection microsurface BRDF~(Equation (15) of \citet{WalterMicrofacet}) becomes
\begin{equation}
  f_r^m(\vecv, \vecl, \vecm) = F(\vecvv, \vecm) \dfrac{\delta(\vecb_r, \vecm) }{4 \left| \vdot{\vecvv}{\vecb_r} \right|^2}.
\end{equation}
We used the fact that $\left| \vdot{\vecl}{\vecb} \right| = \left| \vdot{\vecvv}{\vecb} \right|$.
Similarly, for the transmissive microsurfacet BTDF (Equation~(18) of \citet{WalterMicrofacet}) we get
\begin{equation}
  f_t^m(\vecv, \vecl, \vecm) = \left(1-F(\vecvv, \vecm)\right) \dfrac{\delta(\vecb_t, \vecm) \eta_o^2 }{\left( \eta_i \vdot{\vecv^\prime}{\vecb_t} + \eta_o \vdot{\vecl}{\vecb_t} \right)^2}.
\end{equation}

@@

\paragraph*{Final form.}
The final retro-reflective microfacet model can be written as
\begin{equation} \label{eq:retro_microfacet_model}
  %  f_\text{retro}(\vecv, \vecl) = f(\vecvv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecvv, \vecb)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
  f_\text{retro}(\vecv, \vecl) =
  \begin{cases}
    f_{r, \text{retro}}(\vecv, \vecl) = f_r(\vecvv, \vecl)  & \text{if} \vdot{\vecv}{\vecn}\vdot{\vecl}{\vecn} \geq 0, \\
    f_{t, \text{retro}}(\vecv, \vecl) = f_t(\vecvv, \vecl)  & \text{otherwise}.
  \end{cases}
\end{equation}
In the following section we show in detail that this model is plausible, i.e. that symmetry and energy conservation requirements are fulfilled.
Note that for this we impose one technical assumption on $D$:
we require that the microfacet normal distribution is reflection-symmetric around the normal (which is the case for all isotropic distributions and for anisotropic Phong, Beckmann, and GGX).

@@

  \text{reflect}(\vecll + \vecv, \vecn)

@@

  \text{reflect}(\etal\vecll + \etav\vecv, \vecn)

@@

    \vdot{\vecn}{\vech(\vecll, \vecv)}\\

@@

    -\vdot{\mathbf{t_1}}{\vech(\vecll, \vecv)}\\

@@

As $D$ is reflection-symmetric around the normal, i.e. it takes the form $D(\vech) = D\left(\avdot{\vecn}{\vech}, \lvert \vdot{\mathbf{t_1}}{\vech} \rvert, \lvert \vdot{\mathbf{t_2}}{\vech} \rvert\right)$, we can thus infer that using it with a back-vector obeys symmetry.

@@

    F\left((\vecvv, \vech(\vecvv, \vecl)\right) = F\left(\vdot{\vecvv}{\vech(\vecvv, \vecl)}\right) =

@@

and similarly $G_1(\vecl, \vecvv) = G_1(\vecll, \vecv)$ for the v-cavity masking function.
In summary, we can conclude the symmetry of the BRDF

@@

Analogously, we obtain the symmetry requirement of the BTDF

@@

For the albedo we simply have

@@

Given an implementation of a regular microfacet BSDF, extending it to retro-reflection is straightforward:

@@

\noindent\begin{minipage}[tb]{\linewidth}
\begin{lstlisting}[
    language=GLSL,
    caption=Pseudo code to implement the retro-reflective microfacet model using existing regular microfacet BSDF implementation,
    label=lst:implementation,
    nolol=true,
    frame=trBL
    ]
Sample_result sample_bsdf_retro(vec3 V, vec3 L, vec3 N, Bsdf_params params)
{
    vec3 Vp = -V + 2.0 * dot(V, N) * N;
    return sample_bsdf(V, L, N, params);
}
Eval_result eval_bsdf_retro(vec3 V, vec3 L, vec3 N, Bsdf_params params)
{
    vec3 Vp = -V + 2.0 * dot(V, N) * N;
    return eval_bsdf(V, L, N, params);
}

\end{lstlisting}
\end{minipage}

>>>>

\setciteauthor{Matthias Raab and Laurent Belcour and Frankie Liu and Jamie Portsmouth}
\setcitetitle{The Minimal Retroreflective Microfacet Model (MRM)}
%\setheadtitle{Abbreviated title, only if full title won't fit in page headers}

@@

\lstdefinestyle{snippet}{
  float=tp,
  floatplacement=tbp,
  abovecaptionskip=5pt
}

@@

\teaser{
  \vspace*{-0.5cm}
    \hfill
    \includegraphics[width=0.25\linewidth]{figures/cone/nonretro_cone_cropped_label.png}
    \includegraphics[width=0.25\linewidth]{figures/jacket/nonretro_jacket_cropped_label.png}
    \includegraphics[width=0.25\linewidth]{figures/shoe/nonretro_shoe_cropped_label.png} \\
    \vspace{0.06cm}
    \includegraphics[width=0.25\linewidth]{figures/cone/retro_cone_cropped_label.png}
    \includegraphics[width=0.25\linewidth]{figures/jacket/retro_jacket_cropped_label.png}
    \includegraphics[width=0.25\linewidth]{figures/shoe/retro_shoe_cropped_label.png}
    \hfill
  \caption{Comparison of our ``Minimal Retroreflective Microfacet'' (MRM) retroreflector model (bottom row) obtained by an almost trivial modification of the equivalent roughness GGX microfacet model (top row). The lighting in all cases is aligned with the view direction, exhibiting a strong retro-reflective peak.}
   \vspace{-0.5cm}
  \label{fig:teaser}
}

@@

\begin{abstract}
  \small
  We describe the ``Minimal Retroreflective Microfacet'' (MRM) model, which provides a simple scheme for modifying a pre-existing microfacet BSDF implementation to produce a visually plausible, energy-conserving retroreflective result.
  \end{abstract}

@@

Rendering of retro-reflective materials is useful in a variety of contexts. For safety applications such as road markings, signs, vehicles and high-visibility clothing items, retroreflective materials are heavily used (see Figure~\ref{fig:teaser} for some examples). Such materials are typically designed to be retroreflective via a substructure of elements which preferentially scatter light backwards. A number of CG models have been proposed to model retro-reflective materials, but typically they are either not energy-conserving, not easily integrated into existing rendering systems, or too complex to be practical for production use.

For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least and follows measured observations, b) provably energy-conserving, c) efficient and easy to implement, ideally a small modification to well-understood surface models. We present here a model based on the previously published ``back-vector'' formulation~\cite{BelcourBackvector} which meets these requirements. It is so simple a modification to a regular microfacet model, that we term it the ``Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible retro-reflective result.

\paragraph{Previous works.}
Modeling of retro-reflectivity in the surface scattering function has received little attention so far. For example, \citet{Guo2017Retro} proposed a model to incorporate the retro-reflectivity of prismatic sheets. However, their model only provides pure retro-reflection which would lack artistic control. They later extented this model to account for imperfect retro-reflection with the addition of a microfacet surface (following Beckmann's normal distribution) above the prismatic sheet~\citet{Guo2018Retro}. However, due to the coupling of retro-reflection and rough refraction, they were unable to derive a close form solution for it and relied on fitting of the ground truth model.

@@

We detail the correct usage and derivation of this retro-reflective BSDF model within the microfacet framework~\cite{WalterMicrofacet,HeitzMicrofacet}. We first recall some basic principles from microfacet theory (Section~\ref{sec:microfacets}), then show how this model can be steered from forward- to retro-reflection (Section~\ref{sec:backvector}), and further provide proof that the resulting BSDF is fully plausible (Section~\ref{sec:plausibility}). In Section~\ref{sec:implementation}, we sketch how the model can be implemented in practice, which is done trivially by reusing existing microfacet BSDF implementations. Finally, we provide some visual results and discussion of the behavior (Section~\ref{sec:results}).

@@

\begin{figure}
  \begin{center}
    \begin{tabular} {|p{0.2\textwidth}|p{0.7\textwidth}|}
%    \begin{tabular} {|l|l|}
      \hline
      $\vecv$ & normalized vector to viewer \\
      \hline
      $\vecl$ & normalized vector to light \\
      \hline
      $\vecn$ & surface normal, with the convention that it points to
      the medium with lower index of refraction in cases dealing with $\vecv$ and $\vecl$ pointing to different sides\\
      \hline
      $\mathbf{t_1}, \mathbf{t_2}$ & tangent and bitangent (forming an orthonormal coordinate frame with the normal) \\
      \hline
      $\vechr$ & half vector of reflection, $\vechr = \vechr(\vecv, \vecl) = \frac{\vecv + \vecl}{\lVert \vecv + \vecl \rVert}$ \\
      \hline
      $\vecht$ & half vector of refraction, $\vecht = \vecht(\vecv, \vecl) = -\frac{\left(\etav\vecv + \etal\vecl\right)}{\lVert\etav\vecv + \etal\vecl\rVert}$, where $\etav$ and $\etal$ are the refractive indices of the media $\vecv$ and $\vecl$ point to (always points to medium with lower index of refraction)\\
      \hline
      $\vdot{\vecx}{\vecy}$ & dot product of $\vecx$ and $\vecy$ \\
      \hline
      $\text{reflect}(\mathbf{x}, \mathbf{m})$ & reflection of vector $\mathbf{x}$ on surface with normal $\mathbf{m}$, i.e. $\text{reflect}(\mathbf{x}, \mathbf{m}) = -\mathbf{x} + 2 \vdot{\mathbf{x}}{\mathbf{m}} \mathbf{m}$ \\
%      \hline
%      $\Omega$ & hemisphere of directions around the normal\\
      \hline
      $\Omega, \Omega_{\vecx}$ & hemisphere of directions around the normal, hemisphere of directions around the normal pointing to the side $\vecx$ points to\\
      \hline
    \end{tabular}
  \end{center}
  \caption{
    Notation used.
  }
  \label{fig:notation}
\end{figure}

Using the notation of Figure \ref{fig:notation}, a microfacet BRDF \cite{WalterMicrofacet, HeitzMicrofacet} takes the form (\citet{WalterMicrofacet} Equation (20))
\begin{equation}
  f_r(\vecv, \vecl) = \frac{D(\vechr) \, G_{2}(\vecv,\vecl) \, F(\vechr, \vecv)}{4 \vdot{\vecv}{\vecn} \vdot{\vecl}{\vecn}}
  \label{eq:microfacet_brdf}
\end{equation}
where $D$ is the distribution of microfacets, normalized to fulfill $\int_\Omega D(\vech) \vdot{\vecn}{\vech} d\vech = 1$, $G_{2}$ is the shadowing-masking function, and $F$ is the Fresnel term. The half-vector $\vechr$ is used as the \textit{selector} for microfacets reflecting light from incident direction $\vecl$ to outgoing direction $\vecv$ (\textit{i.e.} only microfacets with normals aligned with the half-vector will reflect light).
Similarly, using the half-vector of refraction as selector, yields the corresponding BTDF
\begin{equation}
  f_t(\vecv, \vecl) =
  \frac{\avdot{\vecv}{\vecht}\avdot{\vecl}{\vecht}\etav^2 D(\vecht) \, G_{2}(\vecv,\vecl) \, \left(1 - F(\vecht, \vecv) \right)}
       {\avdot{\vecv}{\vecn} \avdot{\vecl}{\vecn} \left(\etav \vdot{\vecv}{\vecht} + \etal \vdot{\vecl}{\vecht} \right)^2}
  \label{eq:microfacet_btdf}.
\end{equation}
In Section~4.1 and Section~4.2 of their work, Walter et al.~\cite{WalterMicrofacet} detail the effect of using the half vector as the microfacet selector and obtain Equations~\eqref{eq:microfacet_brdf} and \eqref{eq:microfacet_btdf}.
In this work, we show that we can use the back vector~\cite{BelcourBackvector} used to model retro-reflection as a microfacet selector. Before detailing the impact of such a change, we review some properties of microfacet models.

\paragraph*{Shadowing-Masking.}
The shadowing-masking function combines the visibility masking function $G_1(\vecv, \vecl)$ for view direction and
$G_1(\vecl, \vecv)$ for light direction:
\begin{equation}
G_{2}(\vecv, \vecl) = G_{2}\left(G_1(\vecv, \vecl), G_1(\vecl, \vecv) \right) \ .
\end{equation}
Mathematically, the BSDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
$G_{2}(\vecv, \vecl) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible micro-surfaces \cite{HeitzMicrofacet}
\begin{equation}\label{eq:visible}
  \avdot{\vecv}{\vecn} = \int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \,G_1(\vecv, \text{reflect}(\vecv, \vech)) \vdot{\vecv}{\vech} d\vech.
\end{equation}
Choices for $G_1$ with this property are the Smith masking function
\begin{equation}
  G_1(\vecv, \vecl) = G_{1,\text{Smith}}(\vecv) = \frac{\avdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech},
\end{equation}
where analytic expressions are available for Beckmann and GGX distributions, and the V-cavity masking function
\begin{equation}
  G_1(\vecv, \vecl) = G_{1,\text{vc}}(\vdot{\vecv}{\vecn},\vdot{\vecn}{\vech},\vdot{\vecv}{\vech}) =
  \min\left\lbrace\frac{2\avdot{\vecv}{\vecn}\avdot{\vecn}{\vech}}{\avdot{\vecv}{\vech}}, 1\right\rbrace,
\end{equation}
which is generally applicable for distributions with symmetry around the normal.

@@

As illustrated in Figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as a reflection of the view direction in a ``virtual'' plane aligned with $\vecv$ and perpendicular to the surface.
Intuitively, this is equivalent to observing the microfacet reflection through a virtual mirror, thus turning glossy forward-reflection into glossy retro-reflection.
While there is no immediate physical foundation to this construction, this interpretation can serve as a motivation why the components of the regular microfacet model still work on changing the domain to the back-vector.
We now detail the mathematical consequences of using the back-vector as microfacet selector.

\paragraph*{Jacobian of Back-Vector.}
First we observe that the Jacobian of the change of variable from the back-vector to the light vector $\vecl$ for specular microfacets~(Equation~(11) and~(14) of \citet{WalterMicrofacet}) has the same form as the half-vector one:
\begin{equation}
\left\lVert \dfrac{\partial \omega_{\vecb_{r}}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech_{r}(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{1}{4 \left| \vdot{\vecl}{\vech(\vecvv, \vecl)} \right|} = \dfrac{1}{4 \left| \vdot{\vecl}{\vecb} \right|} \ .
\end{equation}
We also observe that this is true for refractive microfacets~(Equation (17) of \citet{WalterMicrofacet}):
\begin{equation}
\left\lVert \dfrac{\partial \omega_{\vecb_{t}}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech_{t}(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{  \eta_o^2 \left| \vecl, \vech_{t}(\vecvv, \vecl)  \right|}{ \left\lVert \vech_{t}(\vecvv, \vecl) \right\rVert^2}
\end{equation}
This is due to the fact that we have to differentiate with respect to the light direction (which is not altered by the back vector).

\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.99\linewidth]{figures/brdf_plots.pdf}
  \caption{Comparison of the retro-reflective MRM BRDF (blue) to the regular GGX microfacet BRDF (green) and rough-diffuse EON model (orange), as a function of view direction $\theta_o$ for fixed light directions $\theta_i \in {45^\circ, 80^\circ}$. The dashed lines are the roughness $0.1$ case, and the dashed lines roughness $0.666$. The grey horizontal line at $1/\pi$ is the Lambert BRDF.}
  \label{fig:brdf_plots}
\end{figure}

\paragraph*{Microfacet reflection term.}
There is experimental evidence~\cite{BelcourBackvector} that using a Fresnel term based on the back-vector can provide a good match to actual retro-reflective material measurements for a reflection back-vector based BRDF.
As such, we update the formulation of $\rho(\vecv, \vecm)$~(Equation (11) of \citet{WalterMicrofacet}) to use the reflected view vector instead. From the aforementioned equivalence of the Jacobians, it follows that the reflection microsurface BRDF~(Equation (15) of \citet{WalterMicrofacet}) becomes
\begin{equation}
  f_r^m(\vecv, \vecl, \vecm) = F(\vecvv, \vecm) \dfrac{\delta(\vecb_r, \vecm) }{4 \left| \vdot{\vecvv}{\vecb_r} \right|^2}.
\end{equation}
We used the fact that $\left| \vdot{\vecl}{\vecb} \right| = \left| \vdot{\vecvv}{\vecb} \right|$.
Similarly, for the transmissive microsurfacet BTDF (Equation~(18) of \citet{WalterMicrofacet}) we get
\begin{equation}
  f_t^m(\vecv, \vecl, \vecm) = \left(1-F(\vecvv, \vecm)\right) \dfrac{\delta(\vecb_t, \vecm) \eta_o^2 }{\left( \eta_i \vdot{\vecv^\prime}{\vecb_t} + \eta_o \vdot{\vecl}{\vecb_t} \right)^2}.
\end{equation}

@@

\paragraph*{Final form.}
It follows from all this that the final retro-reflective microfacet model can be written simply as
\begin{equation} \label{eq:retro_microfacet_model}
  %  f_\text{retro}(\vecv, \vecl) = f(\vecvv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecvv, \vecb)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
  f_\text{retro}(\vecv, \vecl) =
  \begin{cases}
    f_{r, \text{retro}}(\vecv, \vecl) = f_r(\vecvv, \vecl)  & \text{if} \vdot{\vecv}{\vecn}\vdot{\vecl}{\vecn} \geq 0, \\
    f_{t, \text{retro}}(\vecv, \vecl) = f_t(\vecvv, \vecl)  & \text{otherwise}.
  \end{cases}
\end{equation}
In the following section we show in detail that this model is plausible, i.e. that symmetry and energy conservation requirements are fulfilled.
Note that for this we impose one technical assumption on $D$:
we require that the microfacet normal distribution is reflection-symmetric around the normal (which is the case for all isotropic distributions and for anisotropic Phong, Beckmann, and GGX).

Figure~\ref{fig:brdf_plots} plots the BRDF values of the MRM model (green) compared to the rough diffuse (albedo 1) EON model \cite{Portsmouth2025EON} (orange) and GGX (blue). For both MRM and GGX, the Fresnel factor is set to 1.
The $0.1$ roughness case is a very strong, tight peak (note the logarithmic scale on the y-axis), while the $0.666$ roughness case is a much broader peak.
It is evident that MRM BRDF is the same as GGX,with the peak flipped to be back-scattering (at exactly $\theta_o =-\theta_i$, $\phi_o = \phi_i$).
The $\phi_o = 45^\circ$ case demonstrates the situation when the view direction is not aligned in the same azimuthal plane. In this case. both GGX and MRM are dimmer than Lambert or EON, which compensates for the extra energy of the peaks of MRM and GGX (in which almost all of the energy is concentrated, in the reflection or retro-reflection direction respectively for GGX and MRM).

@@

  \text{reflect}(\vecll + \vecv, \vecn) \ .

@@

  \text{reflect}(\etal\vecll + \etav\vecv, \vecn) \ .

@@

    \vdot{\vecn}{\vech(\vecll, \vecv)}  \ , \\

@@

    -\vdot{\mathbf{t_1}}{\vech(\vecll, \vecv)} \ , \\

@@

Since we assumed $D$ is reflection-symmetric around the normal, i.e. it takes the form
D(\vech) = D\left(\avdot{\vecn}{\vech}, \lvert \vdot{\mathbf{t_1}}{\vech} \rvert, \lvert \vdot{\mathbf{t_2}}{\vech} \rvert\right)
we can thus infer that using it with a back-vector obeys symmetry.

@@

    F\left(\vecvv, \vech(\vecvv, \vecl)\right) = F\left(\vdot{\vecvv}{\vech(\vecvv, \vecl)}\right) =

@@

and similarly $G_1(\vecl, \vecvv) = G_1(\vecll, \vecv)$ for the V-cavity masking function.
In summary, we can conclude the reciprocity symmetry of the BRDF

@@

Analogously, we obtain the reciprocity symmetry requirement of the BTDF

@@

For the directional albedos we simply have

@@

Given an implementation of a regular microfacet BSDF, extending it to retro-reflection is extremely straightforward:

@@

\noindent
\begin{lstlisting}[
    style=snippet,
    language=GLSL,
    caption=Pseudo code to implement the retro-reflective microfacet model using existing regular microfacet BSDF implementation,
    label=lst:implementation,
    nolol=true,
    frame=trBL
    ]
Sample_result sample_bsdf_retro(vec3 V, vec3 L, vec3 N, Bsdf_params params)
{
    vec3 Vp = -V + 2.0 * dot(V, N) * N;
    return sample_bsdf(V, L, N, params);
}
Eval_result eval_bsdf_retro(vec3 V, vec3 L, vec3 N, Bsdf_params params)
{
    vec3 Vp = -V + 2.0 * dot(V, N) * N;
    return eval_bsdf(V, L, N, params);
}
\end{lstlisting}

\section{Results and Discussion}

\label{sec:results}

\paragraph*{Incorporation in a standard microfacet model}

Figure~\ref{fig:conductor_renders} shows renders of a shaderball with either a classic GGX or retro-reflective MRM material applied, here assuming \emph{conducting} (i.e. metallic) microfacets.

\begin{figure}[!b]
    \centering
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/classic_angle_0_cropped.png}
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/classic_angle_45_cropped.png}
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/classic_angle_90_cropped.png} \\
    \vspace{0.06cm}
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/retro_angle_0_cropped.png}
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/retro_angle_45_cropped.png}
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/retro_angle_90_cropped.png}
  \caption{Renders of a metallic shaderball with either a classic GGX (top row) or retro-reflective MRM material (bottom row) applied, for various camera angles and a fixed light direction. The relative angle between the view and light direction is $0^\circ$ (left), $45^\circ$ (middle), and $90^\circ$ (right).}
  \label{fig:conductor_renders}
\end{figure}
